---
title: "Heart_Disease_Analysis"
author: "GROUP 4"
date: "2023-11-02"
output: html_document
---

```{r}
heart <- read.csv("heart_disease_uci.csv", header = T)
head(heart)
```


### 1. Describe the distribution of chest pain types (cp) in patients with potential heart disease.
```{r}
cp_freq <- table(heart$cp)
cp_rel_freq <- cp_freq/sum(cp_freq) * 100

barplot(cp_rel_freq, ylim = c(0, 60), main = "Frequency of Chest Pain Types in People with Potential Heart Disease",
        xlab = "Chest Pain Types", ylab = "Count", )
pie(cp_freq)
```

### 2. Describe the distribution of ecg observation at resting condition (restecg) in patients with potential heart disease.
```{r}
ecg_freq <- table(heart$restecg)
ecg_rel_freq <- ecg_freq/sum(ecg_freq) * 100

barplot(ecg_rel_freq, ylim = c(0, 60), main = "Frequency of Resting ECG Measurements in People with Potential Heart Disease",
        xlab = "ECG", ylab = "Count")
pie(ecg_freq)
```

### 3. Describe the distribution of cholesterol level (chol) in patients with potential heart disease.
```{r}
hist(heart$chol)
boxplot(heart$chol, horizontal = T)
```

### Lots of outliers...

### 4. Is there a correlation between age and maximum heart rate (thalch) in patients?
```{r}
cor(heart$thalch, heart$age, use = "complete.obs")
plot(heart$age, heart$thalch, ylim = c(25, 250), xlim = c(25, 80))
lm(heart$age ~ heart$thalch)
```

### 5. Is there a relationship between cholesterol levels (chol) and geographic regions (origin)?
```{r}
boxplot(heart$chol ~ heart$dataset)
```

### limitations: a lot of 0's for cholesterol levels...

### 6. Are sex and presence of exercise induced angina (exang) independent of each other?
```{r}
table()
```

