---
title: "Heart_Disease_Analysis"
author: "GROUP 4"
date: "2023-11-02"
output: html_document
---

```{r}
heart <- read.csv("heart_disease_uci.csv", header = T)
head(heart)

heart <- heart[,c("age", "sex", "dataset", "cp", "chol", "restecg", "thalch", "exang")]
head(heart)
```


### 1. Describe the distribution of chest pain types (cp) in patients with potential heart disease.
```{r}
cp_freq <- table(heart$cp)
cp_rel_freq <- cp_freq/sum(cp_freq) * 100

barplot(cp_rel_freq, ylim = c(0, 60), xlab = "Chest Pain Types", ylab = "Count",
        main = "Frequency of Chest Pain Types in People with Potential Heart Disease")
pie(cp_freq)
```

### 2. Describe the distribution of ecg observation at resting condition (restecg) in patients with potential heart disease.
```{r}
ecg_freq <- table(heart$restecg)
ecg_rel_freq <- ecg_freq/sum(ecg_freq) * 100

barplot(ecg_rel_freq, ylim = c(0, 60), xlab = "ECG", ylab = "Count", main = "Frequency of Resting ECG Measurements in People with Potential Heart Disease")

pie(ecg_freq)
```

### 3. Describe the distribution of cholesterol level (chol) in patients with potential heart disease.
```{r}
### mean replace missing values in cholesterol levels:
heart$chol[is.na(heart$chol)] <- mean(heart$chol,na.rm=TRUE)

hist(heart$chol)
boxplot(heart$chol, horizontal = T)
```

### Lots of outliers...

### 4. Is there a correlation between age and maximum heart rate (thalch) in patients?
```{r}
# replace missing values with mean value
heart$thalch[is.na(heart$thalch)] <- mean(heart$thalch,na.rm=TRUE)

cor(heart$thalch, heart$age)
plot(heart$age, heart$thalch, ylim = c(25, 250), xlim = c(25, 80),
     xlab = "Age", ylab = "Max Heart Rate", main = "Max Heart Rate by Age")
age_thalch_ls <- lm(heart$thalch ~ heart$age)
summary(age_thalch_ls)
```

The relationship between maximum heart rate (thalch) and age is weak and negative. Maximum heart rate decreases as age increases. 

The correlation coefficient between the variables is -0.35, indicating a low negative correlation between the variables.

The regression model of thalch and age is:

    maximum heart rate = - 0.94281 * age + 187.99649
  
The slope of -0.943 indicates an increase in age would decrease maximum heart rate by 0.943. The intercept of 187.996 represents the maximum heart rate of a new born baby.

### 5. Is there a relationship between cholesterol levels (chol) and geographic regions (origin)?
```{r}
boxplot(heart$chol ~ heart$dataset)
```

### limitations: a lot of 0's for cholesterol levels... all of switzerland's dataset have 0 for cholesterol levels

### 6. Are sex and presence of exercise induced angina (exang) independent of each other?
```{r}
angina_freq <- table(heart$sex, heart$exang)
barplot(angina_freq, legend.text = c("Female", "Male"), col = c("pink", "lightblue"),
        ylab = "Count", xlab = "Exercise Induced Angina", beside = T)
```

### There does not seem to be any association as the distribution for **sex** in the bars is similar for different levels of the **angina** variable.

### H0 - sex and exercise induced angina are independent of each other
### H1 - sex and exercise induced angina are not independent of each other
```{r}
chisq.test(angina_freq)
```

### df = 1; DP = 3.84
### X-squared of 27.582 > DP of 3.84... reject H0 so that there are some association present between the variables **sex** and **exang**.